On discretely entropy conservative discontinuous Galerkin methods

### 2017-07-25 docs/splitform/main.tex
updating paper

\begin{theorem}
\label{lemma:ec1}
The formulation (\ref{eq:ecdg_oneelem})
 %and (\ref{eq:ecdg_oneelem_mat}) are
 is semi-discretely entropy conservative in the sense that
\[
\LRp{\pd{U(\bm{u})}{t},1}_{\widehat{D}} =  \LRa{\psi,1}_{\partial \widehat{D}}-\LRa{{\bm{f}^*, \Pi_N\bm{v}}}_{\partial \widehat{D}}.
\]
\end{theorem}
\begin{proof}
We take the test function $\bm{w}$ to be the projection of the entropy variables.  On the left hand side, we may use the fact that, under a method of lines discretization, 
\[
\bm{u}(x,t) = \sum_{j=1}^{N_p} \bm{u}_j(t) \phi_j(x), \qquad \pd{\bm{u}(x,t)}{t} \in V_h.
\]
Thus, testing with $\Pi_N \bm{v}$ and assuming continuity in time yields 
\[
\LRp{\pd{\bm{u}}{t}, \Pi_N \bm{v}(\bm{u})}_{\widehat{D}} = \LRp{\pd{\bm{u}}{t},\bm{v}(\bm{u})}_{\widehat{D}} = \LRp{\pd{U(\bm{u})}{t},1}_{\widehat{D}}, \qquad \forall \bm{u}\in V_h.
\]

%Let $\bm{v}_q$ denote the entropy variables evaluated in terms of the conservative variables at quadrature points.  Then, the same argument involving the time term can be shown using matrix notation by testing with $\bm{v} = {\bm{P}_q\bm{v}_q}$,
%\[
%\LRp{\bm{P}_q\bm{v}_q}^T\bm{M}\pd{\bm{u}}{t} = \bm{v}_q^T\bm{W}\bm{V}_q\bm{M}^{-1}\bm{M}\pd{\bm{u}}{t} = \bm{1}^T \bm{W} {\rm diag}(\bm{v}_q) \pd{\bm{u}_q}{t} = \bm{1}^T \bm{W} \pd{U(\bm{u}_q)}{t},
%\]
%where we have assumed continuity in time, and have used that $\bm{W}$ is diagonal.  
For the spatial derivative, we expand $\bm{f}_S$ out using Assumption~\ref{eq:assumption} and use Lemma~\ref{lemma:dgd_local}
\begin{align*}
&\LRp{\left.\LRp{D_N \sum_{i=1}^{\infty}\bm{f}_i(\bm{u}_x)\bm{g}_i(\bm{u}_y)}\right|_{y=x}, \Pi_N\bm{v}(x)}_{\widehat{D}} = \sum_{i=1}^{\infty} \LRp{{D_N \bm{f}_i(\bm{u}_x)}, {\bm{g}_i(\bm{u}_x)\Pi_N\bm{v} }}_{\widehat{D}}\\
&= \sum_{i=1}^{\infty}\LRa{\bm{f}_i(\bm{u}_x) \bm{g}_i(\bm{u}_x),\Pi_N\bm{v}}_{\partial \widehat{D}}  - \LRp{ \bm{f}_i(\bm{u}_x), D_N \LRp{\bm{g}_i\Pi_N\bm{v} }}_{\widehat{D}}\\
&= \LRa{f_S(\bm{u}_x,\bm{u}_x), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \Pi_N\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}}\\
&= \LRa{f(\bm{u}_x), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \Pi_N\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}},
\end{align*}
where we have used the consistency of $\bm{f}_S$ in the last step.  
%where $\Pi_N^x, \Pi_N^y$ denote $L^2$ projections taken with respect to the $x$ and $y$ coordinates, respectively.  
The volume terms can be manipulated to yield
\begin{align*}
%&\LRp{\left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}(\Pi_N\bm{v}(x)),\bm{u}(\Pi_N\bm{v}(y)))\LRp{\Pi_N\bm{v}(y)-\Pi_N\bm{v}(x)} }}\right|_{y=x},1}_{\widehat{D}} \\
&\LRp{\left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\LRp{\Pi_N\bm{v}(y)-\Pi_N\bm{v}(x)} }}\right|_{y=x},1}_{\widehat{D}} \\
&= \LRp{ \left.\LRp{D_N\LRp{\psi(y)-\psi(x) }}\right|_{y=x},1}_{\widehat{D}} = \LRp{-D_N \psi(x),1}_{\widehat{D}} = \LRa{-\psi,1}_{\partial \widehat{D}},
\end{align*}
where we have used Definition~\ref{def:tadmor} in going from the first to second integral and Lemma~\ref{lemma:dgd_local} in the final step.  
%\[
%{\rm diag}\LRp{\bm{V}_q\bm{D}_x\bm{P}_q \bm{F}^k_S} \not\in V_h
%\]
%for general flux functions.  However, we may use Lemma~\ref{lemma:sbp} to expand out terms
%\begin{align*}
%\bm{v}^T\bm{V}_q^T\bm{W}{\rm diag}\LRp{\bm{V}_q\bm{D}_x\bm{P}_q \bm{F}^k_S} &= \bm{v}^T\bm{V}_q^T{\rm diag}\LRp{\LRp{\bm{P}_q^T\bm{V}_f^T\bm{W}_f {\rm diag}(\bm{n}_x) \bm{V}_f \bm{P}_q  + \bm{W}\bm{D}_q - \bm{D}_q^T\bm{W} } \bm{F}^k_S}.
%\end{align*}
%Because $\bm{F}^k_S$ is evaluated in terms of the projection of the entropy variables, we may apply Definition~\ref{def:tadmor} to simplify the latter terms
%\begin{align*}
%\bm{v}^T\bm{V}_q^T{\rm diag}\LRp{\LRp{\bm{W}\bm{D}_q - \bm{D}_q^T\bm{W} }\bm{F}^k_S} &=\sum_{i,j=1}^{N_q} \LRp{\bm{V}_q\bm{v}}_i \LRp{\bm{W}\bm{D}_q - \bm{D}_q^T\bm{W} }_{ij}\LRp{\bm{F}^k_S}_{ij}\\
%&=\sum_{i,j=1}^{N_q} \LRp{\bm{W}\bm{D}_q}_{ij} \LRp{\psi_i - \psi_j}.  
%\end{align*}
%
%The spatial term does not 
%\[
%\bm{v}\bm{V}_q^T\bm{W}{\rm diag}\LRp{\bm{V}_q\bm{D}_x\bm{P}_q \bm{F}^k_S}
%\]
\end{proof}

@@

We adapt the construction of local differentiation operators introduced previously to construct a global DG differentiation operator
\begin{equation}
\LRp{D^x_h u,vw}_{\Omega} = \sum_k \LRp{\pd{ \Pi_N u}{x},vw}_{D^k} + \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, vw\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{u - \Pi_Nu },\Pi_N(vw) \bm{n}_x}_{\partial D^k}},
\label{eq:dgd_global}
\end{equation}
where $w(x)$ is a bounded smooth function over each element and $\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{D^x_h u,v}_{D^k}$.  
Entropy conservative DG schemes can be constructed using the fact that $D^x_h$ satisfies a \textit{global} analogue of integration by parts:
\begin{lemma}
\label{lemma:ibp}
Let $D^x_h$ be defined as in (\ref{eq:dgd_global}), and define $v^+ = v$ on $\partial \Omega$.  Then, 
\[
\LRp{D^x_h u,vw}_{\Omega} = \LRa{u,vw\bm{n}_x}_{\partial \Omega} - \LRp{u,D^x_h (vw)}_{\Omega}.
\]
for $u \in H^1(\Oh)$ and $\forall v\in V_h(\Omega_h)$.  
\end{lemma}
\begin{proof}
We may rewrite $\LRp{D^x_h u,vw}_{\Omega}$ using the definition of $D_N$ in (\ref{eq:dgd})
\[
\LRp{D^x_h u,vw}_{\Omega} = \sum_k \LRp{D_N u,vw}_{D^k} + \frac{1}{2}\LRa{\jump{u},vw\bm{n}_x}_{\partial D^k \setminus \partial \Omega}.  
\]
Applying Lemma~\ref{lemma:dgd_local} over each element yields 
\begin{align*}
\LRp{D^x_h u,vw}_{\Omega} &= \sum_k \LRp{-u,D_N (vw)}_{D^k} + \LRa{u + \frac{1}{2}\jump{u},vw\bm{n}_x}_{\partial D^k}\\
&= \sum_k \LRp{-u,D_N (vw)}_{D^k} + \LRa{\avg{u},vw\bm{n}_x}_{\partial D^k}\\
&= \sum_k \LRp{-u,D_N (vw)}_{D^k} - \LRa{\jump{vw},u\bm{n}_x}_{\partial D^k \setminus \partial \Omega} + \LRa{vw,u\bm{n}_x}_{\partial \Omega} \\
&= \LRa{vw,u\bm{n}_x}_{\partial \Omega}-\LRp{u,D^x_h (vw)}_{\Omega}.    
\end{align*}
%Since $\pd{ \Pi_N u}{x}\in V_h$ by assumption, we can integrate by parts the local volume term of (\ref{eq:dgd_global}) on $D^k$.  Using properties of the $L^2$ projection then yields
%\begin{align*}
%\LRp{\pd{ \Pi_N u}{x},v}_{D^k} &= \LRp{\pd{ \Pi_N u}{x}, \Pi_N v}_{D^k} = \LRa{\Pi_N u,\Pi_N v}_{\partial D^k} - \LRp{\Pi_N u, \pd{\Pi_N v}{x}}_{D^k}\\
%&= \LRa{\Pi_N u,\Pi_N v}_{\partial D^k} - \LRp{u, \pd{\Pi_N v}{x}}_{D^k}.  
%\end{align*}
%Substituting this into (\ref{eq:dgd_global}) yields
%\[
%\LRp{D^x_h u,v} = \sum_k  - \LRp{u, \pd{\Pi_N v}{x}}_{D^k} + \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, v\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{u},\Pi_Nv \bm{n}_x}_{\partial D^k}}.
%\]
%For elements which do not intersect with the boundary, the term $\LRa{u^+,v\bm{n}_x}_{\partial D^k}$ can be exchanged with terms from neighboring elements to show
%\[
%\sum_k \frac{1}{2}\LRa{u^+,v\bm{n}_x}_{\partial D^k} = \sum_k \frac{1}{2}\LRa{u,v^+\bm{n}^+_x}_{\partial D^k} = 
%\sum_k - \frac{1}{2}\LRa{u,v^+\bm{n}_x}_{\partial D^k}.  
%\]
%On the boundary $\partial \Omega$, the interface terms satisfy
%\[
%\frac{1}{2}\LRa{u^+,v\bm{n}_x}_{\partial \Omega} = \frac{1}{2}\LRa{u,v\bm{n}_x}_{\partial \Omega} = \LRa{u,v\bm{n}_x}_{\partial \Omega}-\frac{1}{2}\LRa{v^+,u\bm{n}_x}_{\partial \Omega}.
%\]
%Combining these yields the following statement
%\begin{align*}
%\LRp{D^x_h u,v} &= \LRa{u,v\bm{n}_x}_{\partial \Omega} - \LRp{\sum_k  \LRp{\pd{\Pi_N v}{x},u}_{D^k} + \frac{1}{2}\LRa{{v^+ - \Pi_N v}, u\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{v},\Pi_Nu \bm{n}_x}_{\partial D^k}} \\
%&= \LRa{u,v\bm{n}_x}_{\partial \Omega}-\LRp{u,D^x_h v} 
%\end{align*}
%which is the desired analogue of integration by parts.  
\end{proof}
%Integration by parts can also be shown to yield the skew-symmetric form of $\bm{D}^x_h$
%\[
%\LRp{D^x_h u,v} =\frac{1}{2} \sum_k \LRp{\pd{ \Pi_N u}{x},v}_{D^k} - \LRp{u,\pd{\Pi_N v}{x}}_{D^k} + \LRa{\LRp{u_P - \Pi_N u_M}\bm{n}_x, v} + \LRa{{u_M }\bm{n}_x,\Pi_Nv}.
%\]
%Taking $v = u_M$ and cancelling $u_P u_M \bm{n}_x$ terms shows that the differentiation operator is skew-symmetric w.r.t.\ a discrete $L^2$ inner product.

%Lemma~\ref{lemma:ibp} is equivalent to showing that the matrix representation $\bm{D}^x_h$ of the global DG differential operator satisfies a global SBP property with respect to a discrete $L^2$ norm defined by quadrature weights.  
%\begin{lemma}
%Let $\bm{u}$ be degrees of freedom for $u,v \in V_h$.  Define the matrix $\bm{D}^x_h$ such that 
%\[
%\bm{W}
%\]
%\end{lemma}
\begin{remark}
It is possible to define a simpler global DG differentiation operator based on the composition of a discrete differentiation operator and the $L^2$ projection.  Let ${D}^x_h: V_h\rightarrow V_h$ be defined as
\[
%\sum_k \LRp{D^x_h u,v}_{D^k} = \sum_k \LRp{\pd{u}{x},v}_{D^k} + \frac{1}{2}\LRa{\jump{u},v\bm{n}_x}_{\partial D^k}, \qquad \forall v \in V_h.
\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{\pd{u}{x},v}_{D^k} + \frac{1}{2}\LRa{\jump{u},v\bm{n}_x}_{\partial D^k}, \qquad \forall v \in V_h.
\]
Then, using that the range of $D^x_h$ is $V_h$, it is straightforward to show that $D^x_h \Pi_N$ satisfies an analogue of integration by parts, such that
\[
\LRp{D^x_h\Pi_N u, v}_{\Omega} = \LRa{\Pi_N u,v}_{\partial \Omega} - \LRp{D^x_h\Pi_N v, u}_{\Omega}, \qquad \forall u,v\in \L.
\]
%Numerical results using this global differentiation operator result in slightly more accurate results than when using the DG operator (\ref{eq:dgd_global}).  
Unfortunately, while (\ref{eq:dgd_global}) requires only traces of neighboring elements, this construction results in a wide stencil involving all degrees of freedom on neighboring elements due to the presence of $\jump{u}$ and the $L^2$ projection.  This choice of global DG differentiation operator is analogous to the choice of SBP-SAT terms specified in \cite{fernandez2016simultaneous} for SBP operators on simplicial elements without boundary nodes.  
\end{remark}

@@

Extending the entropy stable DG formulation (\ref{eq:ecdg_oneelem}) is done by combining the flux differencing with the global DG differential operator $D^x_h$
\begin{equation}
\sum_{k}\LRp{\pd{\bm{u}}{t},\bm{w}}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}(\bm{u}_x,\bm{u}_y)}\right|_{y=x},\bm{w}}_{D^k} = \LRa{\bm{f}(\bm{u}_x)-\bm{f}^*,\bm{w}}_{\partial \Omega},
%\int_{D^k}\pd{\bm{u}}{t} \bm{w} + \left.\LRp{2D^x_h\bm{f}_S\LRp{\bm{u}(\Pi_N\bm{v}({x})),\bm{u}(\Pi_N\bm{v}({y}))}}\right|_{{y}={x}}\bm{w} + \LRa{\bm{f}^*-\bm{f}(\bm{u}),\bm{w}}_{\partial \Omega}= 0, \qquad \forall \bm{w}\in V_h^n,
\label{eq:ecdg}
\end{equation}
where $\bm{f}^*$ is some numerical flux used in the imposition of boundary conditions, and the global operator $D^x_h$ is understood to be applied to each component of $\bm{f}_S\LRp{\bm{u}(x),\bm{u}(y)}$.

@@

It is also straightforward to show that the formulation (\ref{eq:ecdg}) is locally conservative.  Taking $v = \mathbbm{1}_{D^k}$ in the formulation (\ref{eq:ecdg}) yields
\[
d
\]

\note{It holds but with respect to specifically defined fluxes.  Need to use $v = 1$ on $D^k$ and then symmetry of the global DG operator $D^x_h$.  }

@@

\subsection{On implementation for general flux functions}

@@

We begin by rewriting the global DG differentiation operator $D^x_h$ using the lifting operator  ${L}: V_h^f\rightarrow V_h$
\[
\LRp{{L} u_f,v}_{D^k} = \LRa{u_f,v}_{\partial D^k}, \qquad u_f \in V_h^f, \qquad \forall v\in V_h.
\]
where $V_h^f$ denotes the trace space of the approximation space $V_h$
\[
V_h^f = \LRc{ \left.u\right|_{\partial D^k}, \quad u \in V_h}.
\]
The lift operator can be used to rewrite surface terms which involve the volume $L^2$ projection
\[
\LRa{u,\Pi_Nv}_{\partial D^k} = \LRp{Lu,\Pi_N v}_{D^k} = \LRp{Lu,v}_{D^k}, \qquad u\in V_h, \qquad v\not\in V_h.  
\]
Then, the global DG differentiation operator can be rewritten as 
\begin{equation}
\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{\pd{ \Pi_N u}{x} + \frac{1}{2}L\LRp{u-\Pi_N u}\bm{n}_x,v}_{D^k} + \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, v\bm{n}_x}_{\partial D^k}}.
\label{eq:dgd_globalimplement}
\end{equation}

>>>>

\begin{theorem}
\label{lemma:ec1}
The formulation (\ref{eq:ecdg_oneelem})
 %and (\ref{eq:ecdg_oneelem_mat}) are
 is semi-discretely entropy conservative in the sense that
\[
\LRp{\pd{U(\bm{u})}{t},1}_{\widehat{D}} =  \LRa{\psi,1}_{\partial \widehat{D}}-\LRa{{\bm{f}^*, \Pi_N\bm{v}}}_{\partial \widehat{D}}.
\]
\end{theorem}
\begin{proof}
We take the test function $\bm{w}$ to be the projection of the entropy variables.  On the left hand side, we may use the fact that, under a method of lines discretization, 
\[
\bm{u}(x,t) = \sum_{j=1}^{N_p} \bm{u}_j(t) \phi_j(x), \qquad \pd{\bm{u}(x,t)}{t} \in V_h.
\]
Thus, testing with $\Pi_N \bm{v}$ and assuming continuity in time yields 
\[
\LRp{\pd{\bm{u}}{t}, \Pi_N \bm{v}(\bm{u})}_{\widehat{D}} = \LRp{\pd{\bm{u}}{t},\bm{v}(\bm{u})}_{\widehat{D}} = \LRp{\pd{U(\bm{u})}{t},1}_{\widehat{D}}, \qquad \forall \bm{u}\in V_h.
\]
%Let $\bm{v}_q$ denote the entropy variables evaluated in terms of the conservative variables at quadrature points.  Then, the same argument involving the time term can be shown using matrix notation by testing with $\bm{v} = {\bm{P}_q\bm{v}_q}$,
%\[
%\LRp{\bm{P}_q\bm{v}_q}^T\bm{M}\pd{\bm{u}}{t} = \bm{v}_q^T\bm{W}\bm{V}_q\bm{M}^{-1}\bm{M}\pd{\bm{u}}{t} = \bm{1}^T \bm{W} {\rm diag}(\bm{v}_q) \pd{\bm{u}_q}{t} = \bm{1}^T \bm{W} \pd{U(\bm{u}_q)}{t},
%\]
%where we have assumed continuity in time, and have used that $\bm{W}$ is diagonal.  
For the spatial derivative, we expand $\bm{f}_S$ out using Assumption~\ref{eq:assumption} and use Lemma~\ref{lemma:dgd_local}
\begin{align*}
&\LRp{\left.\LRp{D_N \sum_{i=1}^{\infty}\bm{f}_i(\bm{u}_x)\bm{g}_i(\bm{u}_y)}\right|_{y=x}, \Pi_N\bm{v}(x)}_{\widehat{D}} = \sum_{i=1}^{\infty} \LRp{{D_N \bm{f}_i(\bm{u}_x)}, {\bm{g}_i(\bm{u}_x)\Pi_N\bm{v} }}_{\widehat{D}}\\
&= \sum_{i=1}^{\infty}\LRa{\bm{f}_i(\bm{u}_x) \bm{g}_i(\bm{u}_x),\Pi_N\bm{v}}_{\partial \widehat{D}}  - \LRp{ \bm{f}_i(\bm{u}_x), D_N \LRp{\bm{g}_i\Pi_N\bm{v} }}_{\widehat{D}}\\
&= \LRa{\bm{f}_S(\bm{u}_x,\bm{u}_x), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \Pi_N\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}}\\
&= \LRa{\bm{f}(\bm{u}_x), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \Pi_N\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}},
\end{align*}
where we have used the consistency of $\bm{f}_S$ in the last step.  
%where $\Pi_N^x, \Pi_N^y$ denote $L^2$ projections taken with respect to the $x$ and $y$ coordinates, respectively.  
The volume terms can be manipulated to yield
\begin{align*}
%&\LRp{\left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}(\Pi_N\bm{v}(x)),\bm{u}(\Pi_N\bm{v}(y)))\LRp{\Pi_N\bm{v}(y)-\Pi_N\bm{v}(x)} }}\right|_{y=x},1}_{\widehat{D}} \\
&\LRp{\left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\LRp{\Pi_N\bm{v}(y)-\Pi_N\bm{v}(x)} }}\right|_{y=x},1}_{\widehat{D}} \\
&= \LRp{ \left.\LRp{D_N\LRp{\psi(y)-\psi(x) }}\right|_{y=x},1}_{\widehat{D}} = \LRp{-D_N \psi(x),1}_{\widehat{D}} = \LRa{-\psi,1}_{\partial \widehat{D}},
\end{align*}
where we have used Definition~\ref{def:tadmor} in going from the first to second integral and Lemma~\ref{lemma:dgd_local} in the final step.  
%\[
%{\rm diag}\LRp{\bm{V}_q\bm{D}_x\bm{P}_q \bm{F}^k_S} \not\in V_h
%\]
%for general flux functions.  However, we may use Lemma~\ref{lemma:sbp} to expand out terms
%\begin{align*}
%\bm{v}^T\bm{V}_q^T\bm{W}{\rm diag}\LRp{\bm{V}_q\bm{D}_x\bm{P}_q \bm{F}^k_S} &= \bm{v}^T\bm{V}_q^T{\rm diag}\LRp{\LRp{\bm{P}_q^T\bm{V}_f^T\bm{W}_f {\rm diag}(\bm{n}_x) \bm{V}_f \bm{P}_q  + \bm{W}\bm{D}_q - \bm{D}_q^T\bm{W} } \bm{F}^k_S}.
%\end{align*}
%Because $\bm{F}^k_S$ is evaluated in terms of the projection of the entropy variables, we may apply Definition~\ref{def:tadmor} to simplify the latter terms
%\begin{align*}
%\bm{v}^T\bm{V}_q^T{\rm diag}\LRp{\LRp{\bm{W}\bm{D}_q - \bm{D}_q^T\bm{W} }\bm{F}^k_S} &=\sum_{i,j=1}^{N_q} \LRp{\bm{V}_q\bm{v}}_i \LRp{\bm{W}\bm{D}_q - \bm{D}_q^T\bm{W} }_{ij}\LRp{\bm{F}^k_S}_{ij}\\
%&=\sum_{i,j=1}^{N_q} \LRp{\bm{W}\bm{D}_q}_{ij} \LRp{\psi_i - \psi_j}.  
%\end{align*}
%
%The spatial term does not 
%\[
%\bm{v}\bm{V}_q^T\bm{W}{\rm diag}\LRp{\bm{V}_q\bm{D}_x\bm{P}_q \bm{F}^k_S}
%\]
\end{proof}

@@

We adapt the construction of local differentiation operators introduced previously to construct a global DG differentiation operator
\begin{equation}
\LRp{D^x_h u,vw}_{\Omega} = \sum_k \LRp{\pd{ \Pi_N u}{x},vw}_{D^k} + \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, vw\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{u - \Pi_Nu },\Pi_N(vw) \bm{n}_x}_{\partial D^k}},
\label{eq:dgd_global}
\end{equation}
where $w(x)$ is a bounded smooth function over each element and $\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{D^x_h u,v}_{D^k}$.  
Entropy conservative DG schemes can be constructed using the fact that $D^x_h$ satisfies a \textit{global} analogue of integration by parts:
\begin{lemma}
\label{lemma:ibp}
Let $D^x_h$ be defined as in (\ref{eq:dgd_global}), and define $v^+ = v$ on $\partial \Omega$.  Then, 
\[
\LRp{D^x_h u,vw}_{\Omega} = \LRa{u,vw\bm{n}_x}_{\partial \Omega} - \LRp{u,D^x_h (vw)}_{\Omega}.
\]
for $u \in H^1(\Oh)$ and $\forall v\in V_h(\Omega_h)$.  
\end{lemma}
\begin{proof}
We may rewrite $\LRp{D^x_h u,vw}_{\Omega}$ using the definition of $D_N$ in (\ref{eq:dgd})
\[
\LRp{D^x_h u,vw}_{\Omega} = \sum_k \LRp{D_N u,vw}_{D^k} + \frac{1}{2}\LRa{\jump{u},vw\bm{n}_x}_{\partial D^k \setminus \partial \Omega}.  
\]
Applying Lemma~\ref{lemma:dgd_local} over each element yields 
\begin{align*}
\LRp{D^x_h u,vw}_{\Omega} &= \sum_k \LRp{-u,D_N (vw)}_{D^k} + \LRa{u + \frac{1}{2}\jump{u},vw\bm{n}_x}_{\partial D^k}\\
&= \sum_k \LRp{-u,D_N (vw)}_{D^k} + \LRa{\avg{u},vw\bm{n}_x}_{\partial D^k}\\
&= \sum_k \LRp{-u,D_N (vw)}_{D^k} - \LRa{\jump{vw},u\bm{n}_x}_{\partial D^k \setminus \partial \Omega} + \LRa{vw,u\bm{n}_x}_{\partial \Omega} \\
&= \LRa{vw,u\bm{n}_x}_{\partial \Omega}-\LRp{u,D^x_h (vw)}_{\Omega}.    
\end{align*}
\end{proof}

\begin{remark}
It is possible to define a simpler global DG differentiation operator based on the composition of a discrete differentiation operator and the $L^2$ projection.  Let ${D}^x_h: V_h\rightarrow V_h$ be defined as
\[
%\sum_k \LRp{D^x_h u,v}_{D^k} = \sum_k \LRp{\pd{u}{x},v}_{D^k} + \frac{1}{2}\LRa{\jump{u},v\bm{n}_x}_{\partial D^k}, \qquad \forall v \in V_h.
\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{\pd{u}{x},v}_{D^k} + \frac{1}{2}\LRa{\jump{u},v\bm{n}_x}_{\partial D^k}, \qquad \forall v \in V_h.
\]
Then, using that the range of $D^x_h$ is $V_h$, it is straightforward to show that $D^x_h \Pi_N$ satisfies an analogue of integration by parts, such that
\[
\LRp{D^x_h\Pi_N u, v}_{\Omega} = \LRa{\Pi_N u,v}_{\partial \Omega} - \LRp{D^x_h\Pi_N v, u}_{\Omega}, \qquad \forall u,v\in \L.
\]
%Numerical results using this global differentiation operator result in slightly more accurate results than when using the DG operator (\ref{eq:dgd_global}).  
Unfortunately, while (\ref{eq:dgd_global}) requires only traces of neighboring elements, this construction results in a wide stencil involving all degrees of freedom on neighboring elements due to the presence of $\jump{u}$ and the $L^2$ projection.  This choice of global DG differentiation operator is analogous to the choice of SBP-SAT terms specified in \cite{fernandez2016simultaneous} for SBP operators on simplicial elements without boundary nodes.  
\end{remark}

@@

Extending the entropy stable DG formulation (\ref{eq:ecdg_oneelem}) is done by combining the flux differencing with the global DG differential operator $D^x_h$
\begin{equation}
\sum_{k}\LRp{\pd{\bm{u}}{t},\bm{w}}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},\bm{w}}_{D^k} = \LRa{\bm{f}(\bm{u}_x)-\bm{f}^*,\bm{w}}_{\partial \Omega},
%\int_{D^k}\pd{\bm{u}}{t} \bm{w} + \left.\LRp{2D^x_h\bm{f}_S\LRp{\bm{u}(\Pi_N\bm{v}({x})),\bm{u}(\Pi_N\bm{v}({y}))}}\right|_{{y}={x}}\bm{w} + \LRa{\bm{f}^*-\bm{f}(\bm{u}),\bm{w}}_{\partial \Omega}= 0, \qquad \forall \bm{w}\in V_h^n,
\label{eq:ecdg}
\end{equation}
where $\bm{f}^*$ is some numerical flux used in the imposition of boundary conditions, and the global operator $D^x_h$ is understood to be applied to each component of $\bm{f}_S\LRp{\bm{u}(x),\bm{u}(y)}$.

@@

It is also straightforward to show that the formulation (\ref{eq:ecdg}) is locally conservative.  We follow \cite{chen2017entropy} and take $\bm{w} = \mathbbm{1}_{D^k}$ in the formulation (\ref{eq:ecdg}) yields
\[
\sum_{k}\LRp{\pd{\bm{u}}{t},1}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},1}_{D^k} = \LRa{\bm{f}(\bm{u}_x)-\bm{f}^*,1}_{\partial \Omega \cap D^k}.  
\]
Using Lemma~\ref{lemma:dgd_local} and the symmetry of $\bm{f}_S$ yields that
\[
\LRp{\left.\LRp{D^x_h\bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},1}_{D^k} = \LRa{\bm{f}(\bm{u}_x) ,1}_{\partial D^k} - \LRp{\left.\LRp{D^x_h\bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},1}_{D^k}.
\]
Substituting this back into the first equation cancels volume terms, leaving
\[
\sum_{k}\LRp{\pd{\bm{u}}{t},1}_{D^k} + \LRa{\bm{f}(\bm{u}_x) ,1}_{\partial D^k \setminus \partial \Omega} + \LRa{\bm{f}^*,1}_{\partial \Omega \cap D^k} = 0.  
\]
We note that this definition of local conservation is given in terms of the conservation variables evaluated as a function of the projected entropy variables
\[
\bm{u}_x(x) = \bm{u}\LRp{\LRp{\Pi_N \bm{v}}(x)}.
\]
It is not immediately clear whether this statement of local conservation satisfies the conditions of the classic Lax-Wendroff theorem.  However, this form of local conservation does satisfy a generalized definition of local conservation given in \cite{shi2017local}, and is sufficient to guarantee convergence to the weak solution.

@@

\subsection{On implementation in terms of the flux function}

@@

We begin by rewriting the global DG differentiation operator $D^x_h$ using the lifting operator  ${L}: V_h \rightarrow V_h$ \cite{hesthaven2007nodal, di2011mathematical}
\[
\LRp{{L} u,v}_{D^k} = \LRa{u,v}_{\partial D^k}, \qquad u \in V_h, \qquad \forall v\in V_h.
\]
The lift operator can be used to rewrite surface terms which involve the volume $L^2$ projection
\[
\LRa{u,\Pi_Nv}_{\partial D^k} = \LRp{Lu,\Pi_N v}_{D^k} = \LRp{Lu,v}_{D^k}, \qquad u\in V_h, \qquad v\not\in V_h.  
\]
Then, the global DG differentiation operator can be rewritten as 
\begin{equation}
\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{\pd{ \Pi_N u}{x} + \frac{1}{2}L\LRp{u-\Pi_N u}\bm{n}_x,v}_{D^k} + \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, v\bm{n}_x}_{\partial D^k}}.
\label{eq:dgd_globalimplement}
\end{equation}
